---
title: "Discussion 8: Polynomial Regression"
author: "Christine Cai"
date: "February 23, 2016"
output: html_document
---

***

```{r}
dat = read.table('CH08PR06.txt')
names(dat) = c('y', 'X') # note the capital X
head(dat)
plot(dat$X, dat$y, xlab = 'Age', ylab = 'Steroid Level')
```

Due to the quadratic trend in the data, we may want to try fitting a quadratic model. To fit a polynomial model, we should center the predictor(s) first. Why?
```{r}
x = dat$X - mean(dat$X)
```

Then create the desired higher order terms. Since we want to fit a quadratic model here, we only need $x^2$.
```{r}
x2 = x^2
```

Fit the quadratic model,
```{r}
fit = lm(dat$y ~ x + x2); fit
```

Note that `y` is a column in `dat`, but `x` and `x2` are not.
```{r}
head(dat)
```

That's why the above `lm()` call has different syntax from what we've been doing so far. If you want, you can put `x` and `x2` into dat,
```{r}
dat = cbind(dat, x, x2); head(dat)
```

And fit the model using
```{r}
fit = lm(y ~ x + x2, data = dat); fit
```
