---
title: "Discussion 5"
author: "Christine Cai"
date: "Feb 2, 2016"
output: html_document
---

***

### F Test for Lack of Fit

Read in the data. I'm using 'CH03TA10.txt' for this example; the file can be found using the 'hw datasets' link in Textbook Problems and Datasets on Smartsite.
```{r}
dat = read.table('CH03TA10.txt')
names(dat) = c('y', 'x')
head(dat)
```
What is the test statistic for an F test for lack of fit?

Fit the reduced model.
```{r}
fit.r = lm(y ~ x, data = dat)
```
Fit the full model.
```{r}
fit.f = lm(y ~ 0 + as.factor(x), data = dat)
```
Bring up the ANOVA tables and obtain the test statistic.
```{r}
anova(fit.r)
anova(fit.f)
f.star = (0.012371 - 0.012131)/2/0.0006065; f.star
```
While the above suffices, if you want to utilize R to pull out the numbers for you, so that your results are more precise, you can use `str()` to find where the numbers are stored.
```{r}
str(anova(fit.r))
f.star = 
  (anova(fit.r)$'Sum Sq'[2] - 
     anova(fit.f)$'Sum Sq'[length(anova(fit.f)$'Sum Sq')])/
  (anova(fit.r)$Df[2] - anova(fit.f)$Df[2])/
  anova(fit.f)$'Mean Sq'[2]
f.star
```
If you want R to calculate the test statistic for you,
```{r}
anova(fit.r, fit.f)
```
You need to know the degrees of freedom without looking at the ANOVA tables. The df associated with SSE(R) is $n - 2$. The df associated with SSE(F) is $n - c$, where $c$ is the number of unique levels of $X_i$'s. Here are two ways you can obtain $c$:
```{r}
c = length(unique(dat$x)); c
c = length(levels(as.factor(dat$x))); c
```
How would we find the critical value and p-value?
