---
title: 'Discussion 9: Model Selection'
author: "Christine Cai"
date: "March 1, 2016"
output: html_document
---

***

```{r}
dat = read.table('CH09PR13.txt')
names(dat) = c('y', 'x1', 'x2', 'x3')
head(dat)
```

Suppose the full model we want to consider uses all the first-order terms for the given predictors.
```{r}
fit = lm(y ~ ., data = dat)
summary(fit)
```

### Best Subset

So far, we've only used functions in R's base package. The function we're going to use for best subset regression, `leaps()`, is in the 'leaps' package. First, you'll need to install the package using `install.packages('leaps')`. If you're asked to choose a mirror, pick CA 1.

Installing a package just means downloading some files to your hard drive. So you don't need to do this again the next time you open R.

What you do need to do every time you restart R is to load the package.
```{r}
library('leaps')
```

Ignore the warning.

Read the help file for `leaps()` before attempting to use it. The first argument 'x' is a matrix of predictors; here, we have 'dat' minus its first column.

'y' is a response vector; here, we have the first column of 'dat', or dat$y. 

For your hw problem, you're asked to use $R^2_\mathrm{adj}$, so set `method = 'adjr2'`.
```{r}
fit = leaps(dat[,-1], dat[,1], method = 'adjr2')
fit
```

Suppose we want the 3 best subset regression models.

By default, `leaps()` will return (up to) `nbest = 10` best models with 1 predictor, `nbest = 10` best models with 2 predictors, `nbest = 10` best models with 3 predictors, and so on... These models are stored in `fit$which`; `TRUE` means that variable is included and `FALSE` means it's not.

The respective $R^2_\mathrm{adj}$'s are stored in `fit$adjr2`.

So in order to get the best 3, we need to find out which of the models returned by `leaps()` corresponds to the 3 highest $R^2_\mathrm{adj}$'s.

First, get the orders of the $R^2_\mathrm{adj}$'s from largest to smallest.
```{r}
ind = order(fit$adjr2, decreasing = TRUE)
fit$adjr2
ind
```

The above means that the 4th element of `fit$adjr2` is the largest, the 7th element is the 2nd largest, and so on... So the 3rd element is the smallest.

If we rearrange the order of the elements in `fit$adjr2` based on `ind`, we get `fit$adjr2` in decreasing order.
```{r}
fit$adjr2[ind]
```

So the first 3 are the 3 largest $R^2_\mathrm{adj}$'s.

To get the respective models, we just need to rearrange the rows of `fit$which` based on `ind`.
```{r}
fit$which[ind,]
```

The first 3 are the 3 best models according to $R^2_\mathrm{adj}$. They are:
```{r}
lm(y ~ x1 + x2, data = dat)
lm(y ~ x1 + x2 + x3, data = dat)
lm(y ~ x2, data = dat)
```

***

### Stepwise

The function we'll use for stepwise regression is `?step`. Depending on whether we want forward or backward selection, or the hybrid of the two, the syntax differ.

Suppose we want to do forward selection and consider the range of models from an initial model using just the intercept to the model using all first-order terms for the given predictors.

For `object`, enter the initial model with just the intercept.

For `scope`, enter the formula (without the response) for the model using all first-order terms.

Lastly, the default for `direction` is 'both', so make sure to set `direction = 'forward'`.
```{r}
step(lm(y ~ 1, data = dat), ~ x1 + x2 + x3, direction = 'forward')
```
