---
title: "Discussion 9: Qualitative Predictors"
author: "Christine Cai"
date: "March 1, 2016"
output: html_document
---

***

```{r}
dat = read.table('CH08PR24.txt')
names(dat) = c('y', 'x1', 'x2')
head(dat)
```

As usual, `y` is the response. However, while `x1` is a quantitative predictor, `x2` is a qualitative predictor. Problem 8.24 in your text describes the 2 levels of `x2` as 2 'populations'. Suppose we want to plot the data, while making a clear distinction between the 2 populations.

We want to split `dat` into the groups defined by `x2`. `?split`
```{r}
dat.spl = split(dat, dat$x2); dat.spl
```

`dat.spl` is a list, so we can access its elements with `[[ ]]` or by their names. Let's try plotting `y` vs `x1` for `x2 = 0`.
```{r}
plot(dat.spl[[1]]$x1, dat.spl$'0'$y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)')
```

Or equivalently,
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)'))
```

The `with()` command tells R to look for any objects in our function call (`plot()`) under the environment we give (`dat.spl$'0'`).

We also want to plot `y` vs `x1` for `x2 = 1`. But we want these points to be on the same plot. Recall the function `points()`,
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)'))
with(dat.spl$'1', points(x1, y))
```

Note that we can't tell which points correspond to `x2 = 0` and which correspond to `x2 = 1`. You may choose to distinguish the 2 populations by using different plotting characters, colors, or both. If you don't have access to a color printer, then you must use different plotting characters.
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)'))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
```

The original plotting window was determined by the points corresponding to `x2 = 0`, so some points corresponding to `x2 = 1` may be outside of that window and are not plotted. To ensure that the original plotting window will include all the points, we need to define our own x and y limits in the first `plot()` command. The `xlim` needs to be the range of the `x1` vector in the original `dat`. Similarly, `ylim` needs to be the range of the `y` vector in the original `dat`.
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
```

The last thing we need to do for this plot is to include a legend. `?legend`
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
legend('topleft', legend = c('Non-corner Lots', 'Corner Lots'), 
       pch = 1:2, col = c(1, 'blue'))
```

Just as with a function like `points()`, `legend()` appends to the last plot created.

Now suppose we want to add the estimated regression functions for the 2 populations to our plot. First, fit 2 separate regressions for the 2 populations.
```{r}
fit0 = lm(y ~ x1, data = dat.spl$'0'); fit0
fit1 = lm(y ~ x1, data = dat.spl$'1'); fit1
```

Add the first regression line,
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
legend('topleft', legend = c('Non-corner Lots', 'Corner Lots'), 
       pch = 1:2, col = c(1, 'blue'))
points(dat.spl$'0'$x1, fit0$fit, type = 'l')
```

The line looks a bit strange. A closer look at the dataset will reveal that it is not ordered in terms of `x1`.
```{r}
head(dat)
```

So when R is plotting the points in a line, it is zigzagging back and forth instead of connecting points in order. To fix this, we need the order in which the rows should occur if `x1` is in increasing order. Note that since each row of `dat` is one observation, we cannot just sort `x1`. Instead, we need to sort the rows of `dat` in increasing order of `x1`.
```{r}
ind = order(dat$x1); ind
```

The above means that row 59 of `dat` contains the smallest value for `x1`, followed by row 11, 57, and so on. Row 62 of `dat` contains the largest `x1`. So now we just need to rearrange the rows of `dat` according to the above order we found.
```{r}
dat = dat[ind,]; head(dat)
```

Now we need to rerun our code with this newly sorted `dat`.
```{r}
dat.spl = split(dat, dat$x2); dat.spl
fit0 = lm(y ~ x1, data = dat.spl$'0'); fit0
fit1 = lm(y ~ x1, data = dat.spl$'1'); fit1
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
legend('topleft', legend = c('Non-corner Lots', 'Corner Lots'), 
       pch = 1:2, col = c(1, 'blue'))
points(dat.spl$'0'$x1, fit0$fit, type = 'l')
```

This line looks cleaner than the previous one. Now add the other regression function. You may choose to distinguish the 2 regression functions by using different line types, colors, or both. If you don't have access to a color printer, then you must use different line types.
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
legend('topleft', legend = c('Non-corner Lots', 'Corner Lots'), 
       pch = 1:2, col = c(1, 'blue'))
points(dat.spl$'0'$x1, fit0$fit, type = 'l')
points(dat.spl$'1'$x1, fit1$fit, type = 'l', lty = 2, col = 'blue')
```

Lastly, let's change the `legend()` command to account for the added regression lines.
```{r}
with(dat.spl$'0', plot(x1, y, 
     xlab = 'Assessed Valuation (Thousand Dollars)', 
     ylab = 'Selling Price (Thousand Dollars)', 
     xlim = range(dat$x1), ylim = range(dat$y)))
with(dat.spl$'1', points(x1, y, pch = 2, col = 'blue'))
points(dat.spl$'0'$x1, fit0$fit, type = 'l')
points(dat.spl$'1'$x1, fit1$fit, type = 'l', lty = 2, col = 'blue')
legend('topleft', legend = c('Non-corner Lots', 'Corner Lots'), 
       pch = 1:2, col = c(1, 'blue'), lty = 1:2)
```

***

### Model Fitting

To fit $Y_i = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2 + \varepsilon_i$,
```{r}
fit = lm(y ~ x1*as.factor(x2), data = dat)
summary(fit)
anova(fit)
```
